---
title: "Homework 5"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### \hfill [ECO s394D Summer 2021]{style="float:right"}

##### \hfill [Steven Kim]{style="float:right"}

##### \hfill [sk54923]{style="float:right"}

```{r, include=FALSE}
library(tidyverse)
library(mosaic)
```


### Question 1

#### (A)

$$
se(\hat p) = se(\bar X_{N}) = \sqrt{\frac{p(1-p)}{N}}
$$

#### (B)

$$
\begin{aligned}
& Var(\hat p - \hat q) = Var(\bar X_{N} - \bar Y_{M}) = Var(\bar X_{N}) + Var(\bar Y_{M}) 
= \frac{p(1-p)}{N} + \frac{q(1-q)}{M}
\\
& se(\hat p - \hat q) = \sqrt{\frac{p(1-p)}{N} + \frac{q(1-q)}{M}}
\end{aligned}
$$


#### (C)

$$
\begin{aligned}
& P(event~|~Control~diet) = 97 / (1945+97) = 0.04750245 = \hat p \\
& P(event~|~any~Med~Diet)  = (70+85) / (2097+2030+70+85) = 0.03619804 = \hat q
\end{aligned}
$$

```{r}
predimed = read.csv("predimed.csv", header=TRUE)
xtabs(~group + event, data=predimed)
```

```{r}
control_total = (1945+97)
control_events = 97
p_hat = control_events/control_total
med_total = (2097+2030+70+85)
med_events = 85+70
q_hat = med_events/med_total
```

Plug-in standard error:

$$
\hat {se} (\hat p - \hat q) = \sqrt{\frac{\hat p(1-\hat p)}{N} + \frac{\hat q(1-\hat q)}{M}}
$$

```{r, include=FALSE}
se_plugin = sqrt((p_hat*(1-p_hat))/control_total + (q_hat*(1-q_hat))/med_total)
```

It is `r se_plugin`.


#### (D)

The confidence interval using the plug-in method is $(\hat p - \hat q) ~\pm~ 1.96 \times \hat {se}(\hat p - \hat q)$ = (0.0005145781, 0.0220942425).

```{r}
predimed2 = mutate(predimed,
                  meddiet = ifelse(group == 'Control', 'No', 'Yes'))
```

```{r}
set.seed(90000)
boot1 = do(1000)*{
  # contruct a bootstrapped sample
  predimed_boot = resample(predimed2)
  # recalculate the statistics for the bootstrapped sample
  table1_boot = predimed_boot %>%
    group_by(meddiet) %>%
    summarize(p_event = sum(event=='Yes')/n())
  # return the result of interest
  table1_boot$p_event
}

confint(boot1[,1]-boot1[,2], level=0.95)
```

The confidence interval using the bootstrap method is (0.0003954254, 0.02271635).

They are very close to each other.

#### (E)

To obtain the p-value, construct the z-statistic:
$\frac{(\hat p - \hat q) ~-~ 0}{\hat {se}(\hat p - \hat q)}$ = `r (p_hat - q_hat)/se_plugin`

One-sided p-value is $P(Z > z~statistic~|~H_0) =$ `r pnorm((p_hat - q_hat)/se_plugin, lower.tail = FALSE)`

\newpage

### Question 2

#### (A)

$$
\begin{aligned}
& Var(\hat \Delta) = Var(\bar X_N  - \bar Y_M) = Var(\bar X_N) + Var(\bar Y_M) = \frac{\sigma_X^2}{N} + \frac{\sigma_Y^2}{M}
\\
& se(\hat \Delta) = se(\bar X_N  - \bar Y_M) = \sqrt{\frac{\sigma_X^2}{N} + \frac{\sigma_Y^2}{M}}
\end{aligned}
$$

#### (B)


```{r}
gasprices = read.csv("GasPrices.csv", header=TRUE)
xtabs(~Highway, data=gasprices)

N = 22
M = 79

gasprices_by_highway = gasprices %>%
  group_by(Highway) %>%
  summarize(avg_price = mean(Price), var_price = var(Price))
gasprices_by_highway

Xbar = gasprices_by_highway$avg_price[2]
Ybar = gasprices_by_highway$avg_price[1]
sigma_X = gasprices_by_highway$var_price[2]
sigma_Y = gasprices_by_highway$var_price[1]

se2_plugin = sqrt(sigma_X/N + sigma_Y/M)

c((Xbar - Ybar) - 1.96*se2_plugin, (Xbar - Ybar) + 1.96*se2_plugin)
```

The confidence interval using the plug-in method is (0.009329336, 0.082063069)

```{r}
set.seed(90000)
boot2 = do(1000)*{
  # contruct a bootstrapped sample
  gasprices_boot = resample(gasprices)
  # recalculate the statistics for the bootstrapped sample
  table2_boot = gasprices_boot %>%
    group_by(Highway) %>%
    summarize(avg_price = mean(Price))
  # return the result of interest
  matrix(c(table2_boot$avg_price[2], table2_boot$avg_price[1]), ncol=2)
}

confint(boot2[,1]-boot2[,2], level=0.95)
```

The confidence interval using the bootstrap method is (0.009419492, 0.08384626).

They are very close to each other.

#### (C)

To obtain the p-value, construct the z-statistic:
$\frac{(\hat \Delta) ~-~ 0}{\hat {se}(\hat \Delta)}$ = `r ((Xbar - Ybar) - 0)/se2_plugin`

One-sided p-value is $P(Z > z~statistic~|~H_0) =$ `r pnorm(((Xbar - Ybar) - 0)/se2_plugin, lower.tail = FALSE)`

\newpage

### Question 3

#### (A)

$$
\begin{aligned}
E(X_i) & = \sum_{x=1}^\infty \frac{\lambda^x}{x!}e^{-\lambda}x \\
& = \sum_{x=1}^\infty \frac{\lambda^x}{(x-1)!}e^{-\lambda} \\
& = \sum_{x=0}^\infty \frac{\lambda^{x+1}}{x!}e^{-\lambda} \\
& = \lambda\sum_{x=0}^\infty \frac{\lambda^x}{x!}e^{-\lambda} \\
& = \lambda e^\lambda e^{-\lambda} \\
& = \lambda
\\
E(\bar X_N) & = E(\frac{1}{N}\sum_{i=1}^{N}X_i) \\
& = \frac{1}{N}\sum_{i=1}^{N}E(X_i) \\
& = \frac{1}{N} N \lambda \\
& = \lambda
\end{aligned}
$$

#### (B)

$$
\begin{aligned}
Var(\bar X_N) & = Var(\frac{1}{N}\sum_{i=1}^{N}X_i) \\
& = \frac{1}{N^2}Var(\sum_{i=1}^{N}X_i) \\
& = \frac{1}{N^2}\sum_{i=1}^{N}Var(X_i) \\
& = \frac{1}{N^2} N \lambda = \frac{\lambda}{N} \\
se(\bar X_N) & = \sqrt{\frac{\lambda}{N}}
\end{aligned}
$$

#### (C)

$$
\begin{aligned}
X_i \sim f(X|\lambda) & = \frac{\lambda^x}{x!}e^{-\lambda} \\
L(\lambda) & = \prod_{i=1}^N \frac{\lambda^{x_i}}{x_i!}e^{-\lambda} \\
l(\lambda) &= \sum_{i=1}^N ({x_i} log {\lambda} -log{x_i!} {-\lambda}) \\
\frac{\partial l}{\partial \lambda} & = \sum_{i=1}^N (\frac{1}{\lambda}x_i -1) 
  =\frac{1}{\lambda}\sum_{i=1}^Nx_i - N = 0 \\
\lambda & = \frac{\sum_{i=1}^Nx_i}{N} = \bar X_N
\end{aligned}
$$



#### (D)

make the dataframe: 
```{r}
count = c(12, 18, 15, 8,17, 13, 22, 13, 13, 13, 12, 11, 15, 15, 12,
         8, 20, 12, 14, 11, 9, 15, 16, 20, 9, 15, 13, 19, 18, 14)

maps = data.frame(count)

m_bar = mean(maps$count)
var_m = var(maps$count)
maps_se = sqrt((m_bar / 30))

c(m_bar - 1.96*(maps_se), m_bar + 1.96*(maps_se))
```

The confidence interval using the plug-in method is (12.72455, 15.40879).

```{r}
# bootstrap
set.seed(90000)
boot3 = do(1000)*{
  # contruct a bootstrapped sample
  maps_boot = resample(maps)
  # recalculate the statistics for the bootstrapped sample
  table3_boot = maps_boot %>%
    summarize(avg_count = mean(count))
  # return the result of interest
  table3_boot$avg_count
}

confint(boot3[,1], level=0.95)
```

The confidence interval using the bootstrap method is (12.76667, 15.30083).

Again, they are very close to each other.

\newpage

### Question 4

#### (A)



$$
\begin{aligned}
& \hat F_n(x) = \frac{\sum_{i=1}^n I(X_i \leq x)}{n} \\
& \textrm{Let}~Z_i = I(X_i \leq x). \textrm{Then,}~Z_i~\textrm{takes values either}~0~\textrm{or}~1.\\
& P(Z_i=1) = P(X_i \leq x) = F(x)\\
& \textrm{Therefore},~ Z_i\sim Binomial(F(x))\\
& E(Z_i)=F(x), ~Var(Z_i) = F(x) \times (1-F(x))\\
& E(\hat F_n(x)) = E(\frac{\sum_{i=1}^n Z_i}{n}) = \frac{1}{n} \sum_{i=1}^n E(Z_i) 
  = \frac{1}{n} \sum_{i=1}^n F(x) = \frac{1}{n} \times nF(x) = F(x)\\
& Var(\hat F_n(x)) = Var(\frac{\sum_{i=1}^n Z_i}{n}) = \frac{1}{n^2} \sum_{i=1}^n Var(Z_i) 
  = \frac{1}{n^2} \sum_{i=1}^n F(x)(1-F(x)) \\
  & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~= \frac{1}{n} \times nF(x)(1-F(x))
  = \frac{F(x)(1-F(x))}{n}\\
\end{aligned}
$$


#### (B)



$$
\begin{aligned}
\textrm{For any fixed}~x,~E(\hat F_n(x)) = F(x)~\textrm{and}~Var(\hat F_n(x)) = \frac{F(x)(1-F(x))}{n}.
\end{aligned}
$$

$$
\begin{aligned}
\textrm{By Chebyshev's inequality, for every }& \epsilon > 0, \\
P(|\hat F_n(x) - E(\hat F_n(x))| > \epsilon) &= \frac{Var(\hat F_n(x))}{\epsilon^2} \\
P(|\hat F_n(x) - F_n(x)| > \epsilon) &= \frac{F(x)(1-F(x))}{n\epsilon^2} \\
\textrm{Since} \lim_{n \rightarrow \infty} \frac{F(x)(1-F(x))}{n\epsilon^2} &= 0,\\
P(|\hat F_n(x) - F_n(x)| > \epsilon) \longrightarrow 0 &~ \textrm{as}~n \rightarrow \infty
\end{aligned}
$$



#### (C)

A 95% normal-based confidence interval of the probability that the S&P 500 will lose 10% or more in a given year is (0.05521777, 0.19197324).


```{r}
stocks_bonds = read.csv('stocks_bonds.csv')

cdf=ecdf(stocks_bonds$SP500)
stocks = 89

c(cdf(-0.1) -1.96*sqrt(cdf(-0.1)*(1-cdf(-0.1))/stocks), 
  cdf(-0.1) +1.96*sqrt(cdf(-0.1)*(1-cdf(-0.1))/stocks))
```




